Interoception Scale (Study 2) - Data Cleaning

### 2025-02-14 study2/analysis/1_cleaning.qmd
update

estimate_relation(m, length=50) |> 
  ggplot(aes(x = Duration, y = Predicted)) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2) +
  geom_line() +
  geom_vline(xintercept=5, linetype="dashed", color="red") +
  theme_minimal() +
  ggside::geom_xsidedensity(data=mutate(dfchecks, 
                                        Outlier=ifelse(Outlier==1, "Failed attention check", "Valid")), 
                            aes(fill=Outlier), alpha=0.3) +
  ggside::theme_ggside_void() +
  labs(title = "Experiment Duration",
       subtitle = "Predicted probability of failing attention checks by duration",
       x = "Experiment Duration (minutes)",
       y = "Probability of failing attention checks") 
```

```{r}
outliers$duration <- as.character(df[df$Experiment_Duration < 5, "Participant"])
outliers$duration <- outliers$duration[!outliers$duration %in% outliers$attentionchecks]
```

@@

outliers$distance <- as.character(dfoutlier[dfoutlier$Distance_OPTICS < 4, "Participant"])
outliers$distance <- outliers$distance[!outliers$distance %in% c(outliers$attentionchecks, outliers$duration)]

p1 <- dfoutlier |> 
  ggplot(aes(x=Participant, y=Distance_OPTICS)) +
  geom_bar(aes(fill=Outlier), stat="identity") +
  geom_hline(yintercept = 4, linetype="dashed", color="red") +
  labs(fill = "Status") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.major.x = element_blank())

@@

p2 <- estimate_relation(m, length=30) |> 
  ggplot(aes(x=Distance_OPTICS, y=Predicted)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha=0.2) +
  geom_line() +
  geom_vline(xintercept=4, linetype="dashed", color="red") +
  theme_minimal()

@@

## Imputation

```{r}
#| eval: FALSE

set.seed(3)

d <- datawizard::data_partition(df, proportion = 0.8)

m1 <- glm(MINT_InteroceptiveFailures_10 ~ .,
          family = poisson(),
          data = select(d$p_0.8, starts_with("MINT_"), -contains("AttentionCheck")))

m2 <- parsnip::rand_forest("regression", mtry = 10, trees = 2000) |> 
  fit(MINT_InteroceptiveFailures_10 ~ ., data = select(d$p_0.8, starts_with("MINT_"), -contains("AttentionCheck")))

m3 <- parsnip::mlp(mode="regression") |> 
  fit(MINT_InteroceptiveFailures_10 ~ ., data = select(d$p_0.8, starts_with("MINT_"), -contains("AttentionCheck")))

# m4 <- parsnip::bart(mode="regression") |> 
#   fit(MINT_InteroceptiveFailures_10 ~ ., data = select(d$p_0.8, starts_with("MINT_"), -contains("AttentionCheck")))

library(baguette)

m5 <- parsnip::bag_mlp(mode="regression") |>
  fit(MINT_InteroceptiveFailures_10 ~ ., data = select(d$p_0.8, starts_with("MINT_"), -contains("AttentionCheck")))

# glmmTMB::glmmTMB(MINT_InteroceptiveFailures_10 ~ .,
#                  family = poisson(),
#                  ziformula=~1,
#                  data = select(df, starts_with("MINT_"), -contains("AttentionCheck")))

parsnip::glance(m5)
parsnip::tidy(m5)

results <- data.frame(
  TrueScore = d$test$MINT_InteroceptiveFailures_10,
  # Poisson = predict(m1, d$test),
  RandomForest = predict(m2, d$test)$.pred,
  MLP = predict(m3, d$test)$.pred,
  BagMLP = predict(m5, d$test)$.pred
) |> 
  pivot_longer(-TrueScore)

results |> 
  ggplot(aes(x=TrueScore, y=value, color=name)) +
  geom_abline(slope=1) +
  geom_point()

# Compute R^2
SS_res <- sum((df$TrueScore - df$PredictedScore)^2)
SS_tot <- sum((df$TrueScore - mean(df$TrueScore))^2)
R2 <- 1 - (SS_res / SS_tot)

results |> 
  summarize(RMSE = sqrt(mean((TrueScore - value)^2)),
            SS_res = sum((TrueScore - value)^2),
            SS_tot = sum((TrueScore - mean(value))^2),
            R2 = 1 - (SS_res / SS_tot),
            .by="name")

# Compute R2

```

@@

@@

```{r}
df |> 
  ggplot(aes(x=BMI)) +
  annotate("rect", xmin=10, xmax=18.5, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  annotate("rect", xmin=18.5, xmax=25, ymin=0, ymax=Inf, fill="green", alpha=0.1) +
  annotate("rect", xmin=25, xmax=30, ymin=0, ymax=Inf, fill="yellow", alpha=0.1) +
  annotate("rect", xmin=30, xmax=40, ymin=0, ymax=Inf, fill="orange", alpha=0.1) +
  annotate("rect", xmin=40, xmax=Inf, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  geom_density(aes(color=Gender, fill=Gender), alpha=0.3, linewidth=1) +
  labs(title = "BMI distribution") +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank())

df |> 
  ggplot(aes(x=Age, y=BMI, color=Gender)) +
  geom_point() +
  geom_smooth(aes(fill=Gender), alpha=0.3, method = 'loess', formula = 'y ~ x') +
  labs(title = "BMI and Age") +
  theme_minimal() 
```

### Health

```{r}
df$Disorders_Psychiatric_Mood <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)
df$Disorders_Psychiatric_MoodTreatment <- ifelse(
  df$Disorders_Psychiatric_Mood & !is.na(df$Disorders_PsychiatricTreatment) & str_detect(df$Disorders_PsychiatricTreatment, "Mood|Antidepressant|Anxiolytic|Psychotherapy"), TRUE, FALSE)
# df$Disorders_Psychiatric_Other <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)

p1 <- select(df, Participant, Gender, Disorders_Psychiatric_Mood, Disorders_Psychiatric_MoodTreatment) |> 
  pivot_longer(cols = starts_with("Disorders_"), names_to = "Disorder", values_to = "Value") |> 
  mutate(Disorder = str_remove_all(Disorder, fixed("Disorders_Psychiatric_")),
         Disorder = str_replace(Disorder, "MoodTreatment", "Mood Disorder (with treatment)"),
         Disorder = str_replace(Disorder, "Mood$", "Mood Disorder")) |> 
  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Disorder")) |> 
  mutate(N_tot = sum(N), .by="Disorder") |>
  mutate(Disorder = fct_reorder(Disorder, desc(N_tot))) |>
  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal()

df$Disorders_Somatic_Number <- ifelse(df$Disorders_Somatic == "", 0, str_count(df$Disorders_Somatic, ";")+1)

p2 <- df |> 
  ggplot(aes(x = Disorders_Somatic_Number, fill=Gender)) +
  geom_bar() +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_x_continuous(breaks = c(0:10)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal()

p1 / p2
```

>>>>

estimate_relation(m, length=50) |> 
  ggplot(aes(x = Duration, y = Predicted)) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2) +
  geom_line() +
  geom_vline(xintercept=10, linetype="dashed", color="red") +
  theme_minimal() +
  ggside::geom_xsidedensity(data=mutate(dfchecks, 
                                        Outlier=ifelse(Outlier==1, "Failed attention check", "Valid")), 
                            aes(fill=Outlier), alpha=0.3) +
  ggside::theme_ggside_void() +
  labs(title = "Experiment Duration",
       subtitle = "Predicted probability of failing attention checks by duration",
       x = "Experiment Duration (minutes)",
       y = "Probability of failing attention checks") 
```

```{r}
outliers$duration <- as.character(df[df$Experiment_Duration < 10, "Participant"])
outliers$duration <- outliers$duration[!outliers$duration %in% outliers$attentionchecks]
```

@@

outliers$distance <- as.character(dfoutlier[dfoutlier$Distance_OPTICS < 3, "Participant"])
outliers$distance <- outliers$distance[!outliers$distance %in% c(outliers$attentionchecks, outliers$duration)]

p1 <- dfoutlier |> 
  ggplot(aes(x=Participant, y=Distance_OPTICS)) +
  geom_bar(aes(fill=Outlier), stat="identity") +
  geom_hline(yintercept = 3, linetype="dashed", color="red") +
  labs(fill = "Status") +
  scale_fill_manual(values=c("Duration" = "purple", "Failed Attention Checks"="red", "Passed"="grey")) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.major.x = element_blank())

@@

p2 <- estimate_relation(m, length=30) |> 
  ggplot(aes(x=Distance_OPTICS, y=Predicted)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha=0.2) +
  geom_line() +
  geom_vline(xintercept=3, linetype="dashed", color="red") +
  theme_minimal()

@@

## Dimension Computation

```{r}
#| code-fold: true

compute_and_remove <- function(df, name="BodyAwareness", pattern=name, method="mean") {
  items <- select(df, starts_with(pattern), -contains("AttentionCheck"))
  df <- df[!names(df) %in% names(items)]
  if(method == "mean") {
    df[[name]] <- rowMeans(items, na.rm=TRUE)
  } else {
    df[[name]] <- rowSums(items, na.rm=TRUE)
  }
  df
}
```

df <- compute_and_remove(df, name="IAS", method="sum")

df <- compute_and_remove(df, name="BPQ", pattern="BodyAwareness", method="mean")

df[grepl("^MAIA_.*_R$", names(df))] <- 6 - df[grepl("^MAIA_.*_R$", names(df))]  # Reverse
df <- compute_and_remove(df, name="MAIA_AttentionRegulation", method="mean")
df <- compute_and_remove(df, name="MAIA_BodyListening", method="mean")
df <- compute_and_remove(df, name="MAIA_EmotionalAwareness", method="mean")
df <- compute_and_remove(df, name="MAIA_NotDistracting", method="mean")
df <- compute_and_remove(df, name="MAIA_Noticing", method="mean")
df <- compute_and_remove(df, name="MAIA_NotWorrying", method="mean")
df <- compute_and_remove(df, name="MAIA_SelfRegulation", method="mean")
df <- compute_and_remove(df, name="MAIA_Trusting", method="mean")
```

df <- compute_and_remove(df, name="PHQ4_Anxiety", method="sum")
df <- compute_and_remove(df, name="PHQ4_Depression", method="sum")

# TODO: check if any reversed items in the following questionnaires
df <- compute_and_remove(df, name="TAS_EOT", method="mean")
df <- compute_and_remove(df, name="TAS_DDF", method="mean")
df <- compute_and_remove(df, name="TAS_DIF", method="mean")

df <- compute_and_remove(df, name="ERS_Arousal", method="mean")
df <- compute_and_remove(df, name="ERS_Sensitivity", method="mean")
df <- compute_and_remove(df, name="ERS_Persistence", method="mean")

df <- compute_and_remove(df, name="CERQ_Acceptance", method="mean")
df <- compute_and_remove(df, name="CERQ_PositiveRefocusing", method="mean")
df <- compute_and_remove(df, name="CERQ_PositiveReappraisal", method="mean")
df <- compute_and_remove(df, name="CERQ_SelfBlame", method="mean")
df <- compute_and_remove(df, name="CERQ_OtherBlame", method="mean")
df <- compute_and_remove(df, name="CERQ_Rumination", method="mean")
df <- compute_and_remove(df, name="CERQ_Catastrophizing", method="mean")
df <- compute_and_remove(df, name="CERQ_RefocusPlanning", method="mean")
df <- compute_and_remove(df, name="CERQ_Perspective", method="mean")

df <- compute_and_remove(df, name="CEFSA_Body", method="mean")
df <- compute_and_remove(df, name="CEFSA_Agency", method="mean")
df <- compute_and_remove(df, name="CEFSA_Emotion", method="mean")
df <- compute_and_remove(df, name="CEFSA_Familiarity", method="mean")
df <- compute_and_remove(df, name="CEFSA_Connection", method="mean")
df <- compute_and_remove(df, name="CEFSA_Reality", method="mean")
df <- compute_and_remove(df, name="CEFSA_Self", method="mean")

df[grepl("^PI18.*_R$", names(df))] <- 6 - df[grepl("^PI18.*_R$", names(df))]  # Reverse
df$PI_Alive <- rowMeans(df[grepl("^PI18.*A_", names(df))], na.rm=TRUE)
df <- compute_and_remove(df, name="PI_Safe", pattern="PI18_GS", method="mean")
df <- compute_and_remove(df, name="PI_Enticing", pattern="PI18_GE", method="mean")
df$PI_Good <- rowMeans(select(df, PI_Safe, PI_Enticing), na.rm=TRUE)
df <- select(df, -starts_with("PI18_"))
```

@@

Among these, `r sum(is.na(df$MINT_Awareness_RelA_29))` have missing data for the MINT due to a technical error.

@@

```{r}
p_bmi <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=BMI)) +
  annotate("rect", xmin=10, xmax=18.5, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  annotate("rect", xmin=18.5, xmax=25, ymin=0, ymax=Inf, fill="green", alpha=0.25) +
  annotate("rect", xmin=25, xmax=30, ymin=0, ymax=Inf, fill="yellow", alpha=0.25) +
  annotate("rect", xmin=30, xmax=40, ymin=0, ymax=Inf, fill="orange", alpha=0.25) +
  annotate("rect", xmin=40, xmax=Inf, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  geom_density(aes(color=Gender, fill=Gender), alpha=0.5, linewidth=2) +
  labs(title = "BMI distribution") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank())

p_bmi_a <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=Age, y=BMI, color=Gender)) +
  geom_point(size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender), alpha=0.3, method = 'loess', formula = 'y ~ x') +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none")

p_bmi_b <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=Physical_Active, y=BMI)) +
  geom_jitter(aes(color=Gender), width=0.1, size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender, color=Gender), alpha=0.3, method = 'lm', formula = 'y ~ x') +
  labs(x = "Physically Active", y = "") +
  scale_x_continuous(breaks=0:7) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank()) +
  ggside::geom_xsidehistogram(aes(fill=Gender), bins = 30, position="dodge") +
  ggside::theme_ggside_void()

df$Physical_Workout[df$Physical_Workout > 25] <- NA

p_bmi_c <- df |> 
  filter(!is.na(BMI) & !is.na(Physical_Workout), Gender != "Other") |> 
  ggplot(aes(x=Physical_Workout, y=BMI)) +
  geom_point(aes(color=Gender), size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender, color=Gender), alpha=0.3, method = 'lm', formula = 'y ~ x') +
  labs(x = "Hours of Active Workout", y = "") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none") +
  ggside::geom_xsidehistogram(aes(fill=Gender), bins = 30, position="dodge") +
  ggside::theme_ggside_void()

p_bmi / (p_bmi_a | p_bmi_b | p_bmi_c)
```

### Mental Health

```{r}
df$Disorders_Psychiatric_Mood <- ifelse(str_detect(df$Disorders_Psychiatric, "GAD|Bipolar"), TRUE, FALSE)
df$Disorders_Psychiatric_MoodTreatment <- ifelse(
  df$Disorders_Psychiatric_Mood & !is.na(df$Disorders_PsychiatricTreatment) & str_detect(df$Disorders_PsychiatricTreatment, "Mood|Antidepressant|Anxiolytic|Psychotherapy"), TRUE, FALSE)

df$Disorders_Psychiatric_Anxiety <- ifelse(str_detect(df$Disorders_Psychiatric, "Panic|GAD|Social Phobia|Phobia"), TRUE, FALSE)
df$Disorders_Psychiatric_AnxietyTreatment <- ifelse(
  df$Disorders_Psychiatric_Anxiety & !is.na(df$Disorders_PsychiatricTreatment) & str_detect(df$Disorders_PsychiatricTreatment, "Anxiolytic|Psychotherapy|Mindfulness"), TRUE, FALSE)

df$Disorders_Psychiatric_Eating <- ifelse(str_detect(df$Disorders_Psychiatric, "Eating"), TRUE, FALSE)
df$Disorders_Psychiatric_Addiction <- ifelse(str_detect(df$Disorders_Psychiatric, "Addiction"), TRUE, FALSE)
df$Disorders_Psychiatric_Borderline <- ifelse(str_detect(df$Disorders_Psychiatric, "BPD"), TRUE, FALSE)

df$Disorders_Psychiatric_Autism <- ifelse(str_detect(df$Disorders_Psychiatric, "ASD"), TRUE, FALSE)
df$Disorders_Psychiatric_ADHD <- ifelse(str_detect(df$Disorders_Psychiatric, "ADHD"), TRUE, FALSE)

# df$Disorders_Psychiatric_Other <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)

select(df, Participant, Gender, 
             Disorders_Psychiatric_Mood, Disorders_Psychiatric_MoodTreatment,
             Disorders_Psychiatric_Anxiety, Disorders_Psychiatric_AnxietyTreatment,
             Disorders_Psychiatric_Eating, Disorders_Psychiatric_Addiction, Disorders_Psychiatric_Borderline,
             Disorders_Psychiatric_Autism, Disorders_Psychiatric_ADHD) |> 
  pivot_longer(cols = starts_with("Disorders_"), names_to = "Disorder", values_to = "Value") |> 
  mutate(Disorder = str_remove_all(Disorder, fixed("Disorders_Psychiatric_")),
         Disorder = str_replace(Disorder, "MoodTreatment", "Mood Disorder (with treatment)"),
         Disorder = str_replace(Disorder, "AnxietyTreatment", "Anxiety (with treatment)"),
         Disorder = str_replace(Disorder, "Mood$", "Mood Disorder")) |> 
  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Disorder")) |> 
  mutate(N_tot = sum(N), .by="Disorder") |>
  mutate(Disorder = fct_reorder(Disorder, desc(N_tot)),
         Treatment = ifelse(str_detect(Disorder,  "Treatment"), "With treatment", "Without treatment")) |>
  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

### Somatic

```{r}
somatic <- data.frame(
  Hypermobility = ifelse(str_detect(df$Disorders_Somatic, "Hypermobility Syndrome"), TRUE, FALSE),
  Fibromyalgia = ifelse(str_detect(df$Disorders_Somatic, "Fibromyalgia"), TRUE, FALSE),
  ChronicFatigue = ifelse(str_detect(df$Disorders_Somatic, "Chronic Fatigue Syndrome"), TRUE, FALSE),
  ChronicPain = ifelse(str_detect(df$Disorders_Somatic, "Chronic Pain Syndrome"), TRUE, FALSE),
  BackPain = ifelse(str_detect(df$Disorders_Somatic, "Back Pain"), TRUE, FALSE),
  MuscleTension = ifelse(str_detect(df$Disorders_Somatic, "Muscle Tension"), TRUE, FALSE),
  SkinRashes = ifelse(str_detect(df$Disorders_Somatic, "Skin Rashes"), TRUE, FALSE),
  Eczema = ifelse(str_detect(df$Disorders_Somatic, "Eczema"), TRUE, FALSE),
  Psoriasis = ifelse(str_detect(df$Disorders_Somatic, "Psoriasis"), TRUE, FALSE),
   # Dermatological = ifelse(str_detect(df$Disorders_Somatic, "Other Dermatological"), TRUE, FALSE),
  # Musculoskeletal = ifelse(str_detect(df$Disorders_Somatic, "Other Musculoskeletal"), TRUE, FALSE),
  Sjogrens = ifelse(str_detect(df$Disorders_Somatic, "Sjogren"), TRUE, FALSE),
  ChestPain = ifelse(str_detect(df$Disorders_Somatic, "Chest Pain"), TRUE, FALSE),
  Arrhythmia = ifelse(str_detect(df$Disorders_Somatic, "Cardiac Arrhythmia"), TRUE, FALSE),
  Hypertension = ifelse(str_detect(df$Disorders_Somatic, "Hypertension"), TRUE, FALSE),
  Hypotension = ifelse(str_detect(df$Disorders_Somatic, "Hypotension"), TRUE, FALSE),
  IBS = ifelse(str_detect(df$Disorders_Somatic, "IBS"), TRUE, FALSE),
  GERD = ifelse(str_detect(df$Disorders_Somatic, "GERD"), TRUE, FALSE),
  Crohn = ifelse(str_detect(df$Disorders_Somatic, "Crohn"), TRUE, FALSE),
  UlcerativeColitis = ifelse(str_detect(df$Disorders_Somatic, "Ulcerative Colitis"), TRUE, FALSE),
  Celiac = ifelse(str_detect(df$Disorders_Somatic, "Celiac Disease"), TRUE, FALSE),
  Gluten = ifelse(str_detect(df$Disorders_Somatic, "Gluten"), TRUE, FALSE),
  Lactose = ifelse(str_detect(df$Disorders_Somatic, "Lactose"), TRUE, FALSE),
  ShortBreath = ifelse(str_detect(df$Disorders_Somatic, "Shortness of Breath"), TRUE, FALSE),
  Asthma = ifelse(str_detect(df$Disorders_Somatic, "Asthma"), TRUE, FALSE),
  COPD = ifelse(str_detect(df$Disorders_Somatic, "COPD"), TRUE, FALSE),
  SleepApnea = ifelse(str_detect(df$Disorders_Somatic, "Sleep Apnea"), TRUE, FALSE),
  Bronchitis = ifelse(str_detect(df$Disorders_Somatic, "Chronic Bronchitis"), TRUE, FALSE),
  Nausea = ifelse(str_detect(df$Disorders_Somatic, "Nausea/Vomiting"), TRUE, FALSE),
  Dizziness = ifelse(str_detect(df$Disorders_Somatic, "Dizziness/Lightheadedness"), TRUE, FALSE),
  Migraine = ifelse(str_detect(df$Disorders_Somatic, "Migraine"), TRUE, FALSE),
  Neuropathy = ifelse(str_detect(df$Disorders_Somatic, "Neuropathy"), TRUE, FALSE),
  # Epilepsy = ifelse(str_detect(df$Disorders_Somatic, "Epilepsy"), TRUE, FALSE),
  MS = ifelse(str_detect(df$Disorders_Somatic, "Multiple Sclerosis"), TRUE, FALSE),
  FrequentUrination = ifelse(str_detect(df$Disorders_Somatic, "Frequent Urination"), TRUE, FALSE),
  Endometriosis = ifelse(str_detect(df$Disorders_Somatic, "Endometriosis"), TRUE, FALSE),
  Cystitis = ifelse(str_detect(df$Disorders_Somatic, "Interstitial Cystitis"), TRUE, FALSE),
  PelvicPain = ifelse(str_detect(df$Disorders_Somatic, "Chronic Pelvic Pain Syndrome"), TRUE, FALSE)
)

somatic |> 
  select(where(~n_distinct(.) > 1)) |> 
  correlation::correlation(redundant = TRUE, p_adjust="none") |>
  correlation::cor_sort() |> 
  mutate(label = ifelse(p < .001, "*", ""),
         label = ifelse(r == 1, "", label)) |> 
  ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
  geom_tile() +
  geom_text(aes(label=label), size=3) +
  scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
  labs(title = "Co-occurence Matrix of Somatic Disorders") +
  theme_minimal()  +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_blank())

# pca <- as.data.frame(prcomp(somatic)$rotation[, 1:5])
# pca$PC <- colnames(pca)[max.col(pca, ties.method='first')]
# 
# pca |> 
#   rownames_to_column(var="Disorder") |> 
#   pivot_longer(-c(Disorder, PC)) |> 
#   filter(PC == name) |> 
#   arrange(name, desc(value)) |> 
#   ggplot(aes(x=value, y=Disorder, fill=name)) +
#   geom_bar(stat = "identity") +
#   facet_wrap(~name, scale="free_y")
```

```{r}
#| fig-keep: "first"

# somatic |> 
#   select(where(~n_distinct(.) > 1)) |> 
#   EGAnet::bootEGA(
#     corr="pearson",
#     EGA.type = "EGA.fit",
#     model = "glasso", 
#     algorithm = "leiden",
#     seed=42)
```

```{r}
# df$Disorders_Somatic_Number <- ifelse(df$Disorders_Somatic == "", 0, str_count(df$Disorders_Somatic, ";")+1)

# 1. Autonomic Dysfunction & Anxiety-Somatic Symptoms
# Mechanism: Dysregulated autonomic nervous system (e.g., panic disorder, dysautonomia, hyperventilation).
df$Disorders_Somatic_Autonomic_N <- somatic |> 
  select(Dizziness, ChestPain, ShortBreath, Arrhythmia, Nausea, Hypotension, Hypertension, MuscleTension) |> 
  rowSums()
df$Disorders_Somatic_Autonomic <- ifelse(df$Disorders_Somatic_Autonomic_N >= 1, TRUE, FALSE)

# 2. Functional Somatic Syndromes (Pain, Fatigue, GI Symptoms, and Anxiety)
# Mechanism: Central sensitization, stress-induced HPA axis dysregulation, visceral hypersensitivity.
df$Disorders_Somatic_Functional_N <- somatic |> 
  select(Fibromyalgia, Neuropathy, Migraine, ChronicFatigue, ChronicPain, IBS, BackPain, PelvicPain, FrequentUrination) |> 
  rowSums()
df$Disorders_Somatic_Functional <- ifelse(df$Disorders_Somatic_Functional_N >= 1, TRUE, FALSE)

# 3. Hypervigilance, Mast Cell Dysfunction & Multi-System Reactivity
# Mechanism: Mast cell activation, histamine intolerance, neuroinflammation, heightened interoception.
df$Disorders_Somatic_Hyperreactivity_N <- somatic |> 
  select(Migraine, SkinRashes, Nausea, FrequentUrination, Dizziness, MuscleTension, Endometriosis) |> 
  rowSums()
df$Disorders_Somatic_Hyperreactivity <- ifelse(df$Disorders_Somatic_Hyperreactivity >= 1, TRUE, FALSE)

# 4. Immune-Inflammatory & Gut-Brain Axis Dysregulation
# Conditions: Celiac disease, gluten sensitivity, lactose intolerance, MS, neuropathy, eczema, asthma, IBS.
# Mechanism: Leaky gut, immune activation, microbiome imbalance, chronic low-grade inflammation.
df$Disorders_Somatic_Immune_N <- somatic |> 
  select(Celiac, Gluten, Lactose, MS, Eczema, Asthma, IBS) |> 
  rowSums()
df$Disorders_Somatic_Immune <- ifelse(df$Disorders_Somatic_Immune_N >= 1, TRUE, FALSE)

# 5. Cardiopulmonary & Stress-Linked Disorders
# Mechanism: Vascular and respiratory dysregulation, autonomic instability, stress-exacerbated cardiovascular symptoms.
df$Disorders_Somatic_Cardio_N <- somatic |> 
  select(Arrhythmia, COPD, ChestPain, ShortBreath, Hypertension, SleepApnea) |> 
  rowSums()
df$Disorders_Somatic_Cardio <- ifelse(df$Disorders_Somatic_Cardio_N >= 1, TRUE, FALSE)

# 6. Connective Tissue & Dysautonomia-Linked Disorders
# Conditions: Hypermobility, Sjogren’s syndrome, hypotension, chronic fatigue, dizziness.
# Mechanism: Connective tissue laxity affecting autonomic function (e.g., Ehlers-Danlos syndrome, POTS).
df$Disorders_Somatic_Connective_N <- somatic |> 
  select(Hypermobility, Hypotension, Sjogrens, Hypotension, ChronicFatigue, Dizziness) |> 
  rowSums()
df$Disorders_Somatic_Connective <- ifelse(df$Disorders_Somatic_Connective_N >= 1, TRUE, FALSE)

p2 <- df |> 
  select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
  select(ends_with("_N"), Gender, Participant) |>
  pivot_longer(starts_with("Disorders_Somatic_")) |> 
  summarize(n = n(), .by=c("name", "value", "Gender")) |> 
  mutate(name = str_remove(name, "Disorders_Somatic_"),
         name = str_remove(name, "_N")) |> 
  ggplot(aes(x = value, y=n, fill=Gender)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Number of Issues") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~name)
p2
# 
# p2 <- df |> 
#   select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
#   select(-ends_with("_N"), Gender, Participant) |>
#   pivot_longer(starts_with("Disorders_Somatic_")) |> 
#   summarize(n = n(), .by=c("name", "value", "Gender")) |> 
#   filter(value == TRUE) |> 
#   ggplot(aes(x = name, y=n, fill=Gender)) +
#   geom_bar(stat="identity") +
#   scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
#   scale_y_continuous(expand = c(0, 0)) +
#   labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Type") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
# p2
```

### Wearables

```{r}
#| fig-height: 12

p1 <- df |> 
  ggplot(aes(x = Wearables_Number)) +
  geom_bar() +
  scale_x_continuous(breaks = c(0:max(df$Wearables_Number)), expand=c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Number of Wearable Devices", y = "Number of Participants", x="Number Owned") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal()

p2 <- df |> 
  select(contains("Wearables"), -contains("Importance"), -Wearables_Number) |> 
  pivot_longer(everything()) |> 
  summarize(n = n(), .by=c("name", "value")) |> 
  mutate(value = fct_relevel(value, "Not owning", "Never", "Very rarely", "A few times per week", "Once a day", "A few times per day", "A few times per hour"),
         name = str_remove(name, "Wearables_")) |> 
  ggplot(aes(x=value, y=n, fill=name)) +
  geom_bar(stat = "identity") +
  labs(x = "Usage frequency", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

get_databubble <- function(df, var="Heart") {
  x <- paste0("Wearables_", var)
  y <- paste0("Wearables_", var, "Importance")
  dat <- df[df[[x]] != "Not owning", ]
  dat <- summarize(dat, n = n(), .by = all_of(c(x, y)))
  lvl <- c("Never", "Very rarely", "A few times per week", "Once a day", "A few times per day", "A few times per hour")
  dat[[x]] <- factor(dat[[x]], levels=lvl[lvl %in% unique(dat[[x]])])
  names(dat) <- c("Usage", "Importance", "n")
  dat$Device <- var
  dat
}
